\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-10-07 chapters/bci.tex
All text

\todo{comments liesel paradigms}
\todo{spellcheck}
\todo{are all acronyms defined}
\todo{fix all latex warnings to make sure no refs are missing (or grep on that
specific warning)}
\todo{uniform ref or autoref}
\todo{don't capitalize figure, table, algorithm,...}
\todo{define short captions for tables and figures}
\todo{fix lacheck/chktex notifications}
\todo{use terminology individuals with severe
speech and physical impairments (SSPI), avoid phrasing 'suffering from'}
\todohsm{more subtitles}
\todofc{double check statements about what has been published 'for the sake of clarity, the introduction has been mofed to"}
\todofc{chapters should look more like full papers}
\todo{uniform vsa condition vsa setting}
\todo{acronyms written in full every new chapter and in captions}
\todo{search for ??}
\todo{acro usage in chapters}
\chapter{Brain-computer interfaces}
\todo{are textcites used when necessary?}
\label{sec:bci}
People with \ac{sspi} due to
neurodegenenerative disease, stroke, traumatic brain or spine injury
become entirely reliant on caregivers or family members in their day to day
routine.
Wouldn't it be great if we could help them regain some quality of life by
making them more independent?

@@

\fullpagefig[topleft]{%
  \includegraphics[width=\fullpagefigwidth]{figures/bci/bci_loop.pdf}
}{%
  The \ac{bci} loop.
}{%
  The \ac{bci} loop.
  The user interacts with the \ac{bci} via a specific paradigm, in this case involving
  visual stimulation.
  Electrical neural signals are recorded, and neurophysiological features
  related to the paradigm are extracted. Using machine learning, a decision can
  then be made based on these features, which can be presented back to the user.
  In a closed-loop \ac{bci}, this feedback allows the user to adapt to the \ac{bci}.
}{%
  fig:bci/loop
}

@@

\fullpagefig{%
  \input{figures/bci/paradigms.tikz.tex}
}{%
  Overview of \ac{bci} paradigms.
}{%
  A \ac{bci} paradigm defines which brain activity will be used for control.
  They can be categorized in active, reactive and passive
  paradigms, depending on their reliance on external stimulation and active
  participation required from the user.
  Adapted from~\textcite{Muehl2014}.
}{%
  fig:bci/paradigms
}

@@

\begin{figure}[t]
  \begin{subfigure}[t]{.45\textwidth}
  \includegraphics[width=\textwidth, height=\textwidth]{example-image-a}
  \end{subfigure}\hfill%
  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/bci/p300.tikz.tex}
    \caption{The P300 matrix speller interface. Entire rows or columns flash at once.}
  \end{subfigure}

  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/bci/erp.tikz.tex}
    \caption{The idealized \ac{erp} evoked by an attended stimulation.}
  \end{subfigure}\hfill%
  \begin{subfigure}[t]{.45\textwidth}
    \includegraphics[width=\textwidth, height=0.6180\textwidth]{example-image-b}
  \end{subfigure}
\end{figure}

@@

\subsection{The \acf{erp}}%
\label{sec:bci/oddball/erp}

>>>>

\todo{comments liesel paradigms}
\todo{spellcheck}
\todo{are all acronyms defined}
\todo{fix all latex warnings to make sure no refs are missing (or grep on that
specific warning)}
\todo{uniform ref or autoref}
\todo{don't capitalize figure, table, algorithm,...}
\todo{define short captions for tables and figures}
\todo{fix lacheck/chktex notifications}
\todo{use terminology individuals with severe
speech and physical impairments (SSPI), avoid phrasing 'suffering from'}
\todohsm{more subtitles}
\todofc{double check statements about what has been published 'for the sake of clarity, the introduction has been mofed to"}
\todofc{chapters should look more like full papers}
\todo{uniform vsa condition vsa setting}
\todo{acronyms written in full every new chapter and in captions}
\todo{search for ??}
\todo{acro usage in chapters}

\chapter{Brain-computer interfaces}
\todo{are textcites used when necessary?}
\label{sec:bci}
People with \ac{sspi} due to
neurodegenenerative disease, stroke, traumatic brain or spine injury
become entirely reliant on caregivers or family members in their day to day
routine.
Wouldn't it be great if we could help them regain some quality of life by
making them more independent?

@@

\fullpagefig[topleft]{%
  \includegraphics[width=\fullpagefigwidth]{figures/bci/bci_loop.pdf}
}{%
  The \acs{bci} loop.
}{%
  The \ac{bci} loop.
  The user interacts with the \ac{bci} via a specific paradigm, in this case involving
  visual stimulation.
  Electrical neural signals are recorded, and neurophysiological features
  related to the paradigm are extracted. Using machine learning, a decision can
  then be made based on these features, which can be presented back to the user.
  In a closed-loop \ac{bci}, this feedback allows the user to adapt to the \ac{bci}.
}{%
  fig:bci/loop
}

@@

\fullpagefig{%
  \input{figures/bci/paradigms.tikz.tex}
}{%
  Overview of \acs{bci} paradigms.
}{%
  A \ac{bci} paradigm defines which brain activity will be used for control.
  They can be categorized in active, reactive and passive
  paradigms, depending on their reliance on external stimulation and active
  participation required from the user.
  Adapted from~\textcite{Muehl2014}.
}{%
  fig:bci/paradigms
}

@@

\begin{figure}[t]
  \begin{subfigure}[t]{.45\textwidth}
  \includegraphics[width=\textwidth,
    height=\textwidth]{figures/bci/illustration}
    \caption{Stimuli are flashed while the \ac{eeg} is recorded}
  \end{subfigure}\hfill%
  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/bci/p300.tikz.tex}
    \caption{The P300 matrix speller interface. Entire rows or columns flash at once.}
  \end{subfigure}

  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/bci/erp.tikz.tex}
    \caption{The idealized \ac{erp} evoked by an attended stimulation in a
    single channel.}
  \end{subfigure}\hfill%
  \begin{subfigure}[t]{.45\textwidth}
    \includegraphics[width=\textwidth, height=0.6180\textwidth]{example-image-b}
  \end{subfigure}
\end{figure}

@@

\subsection{The \acl{erp}}%
\label{sec:bci/oddball/erp}

### 2024-10-07 chapters/bttda.tex
funding
All text

\subsection{Higher Order Discriminant Analysis}
Higher Order Discriminant Analysis (HODA)~\cite{Phan2010} is a
supervised tensor-based feature extraction technique.
For a set of $N$ tensors of order $K$
$\left\{\ten{X}(n)\in\mathbb{R}^{D_1\times D_2 \times \cdots \times
		D_K}\right\}_n^N$, HODA finds projection matrices $\mat{U_k}$ for each mode $k$
that project a given $\ten{X}$ to a latent tensor
$\ten{G}\in\mathbb{R}^{r_1\times r_2\times\cdots\times r_K}$, usually with lower
dimensionality $(r_1\leq D_1,r_2\leq D_2,\ldots,r_K\leq D_K)$ using
tensor-matrix mode products
\begin{equation}
	\ten{G}  = \ten{X}\mmpr{\mat{U}}
	\label{eq:HODA-backward}
\end{equation}
visualized in Figure~\ref{fig:HODA-backward}.
\begin{figure}[t]
	\centering
	\input{figures/bttda/HODA_backward.tikz.tex}
  \caption[The \ac{hoda} backward projection.]{%
    A visualization of the multilinear projection obtained by Higher Order
    Discriminant Analysis (HODA) applied to a third-order tensor
    sample $\ten{X}$ with shape $(D_1,D_2, D_3)$.
		HODA finds projection matrices $\mat{U}_k$ such that maximal
		discriminability between classes is achieved in the projected latent tensors
		$\ten{G}$ with reduced dimensionality $(r_1,r_2,r_3)$.}
	\label{fig:HODA-backward}
\end{figure}
Analogous to the HOSVD, HODA is a dimensionality
reduction decomposition that results in a dense latent tensor $\ten{G}$, and
imposes an orthogonality constraint on $\mat{U}_k$ to ensure uniqueness.
However, while for the HOSVD decomposition the projection matrices
are chosen to minimize the reconstruction error, the projection matrices
$\mat{U}_k$ of HODA are optimized for maximal discriminability of tensors
$\ten{G}(n)$ belonging to classes with labels $c_n$.
This is a desirable property in a classification setting where samples are
high-dimensional tensors.

HODA optimizes discriminability in the Fisher sense, by optimizing the
Fisher ratio $\phi$ between the latent tensors $\ten{G}(n)$
\begin{equation}
	\phi = \frac{\sum_c^CN_c\left\lVert\bar{\ten{G}}(c)-\bar{\bar{\ten{G}}}\right\rVert_F^2}
	{\sum_n^N\left\lVert\ten{G}(n)-\bar{\ten{G}}(c_n)\right\rVert_F^2}
	\label{eq:fisher}
\end{equation}
for $C$ classes with each $N_c$ samples. $\bar{\ten{G}}(c)$ is the mean of
latent
tensors of class $c$, and $\bar{\bar{\ten{G}}}$ the mean of
these class mean latent tensors.
If the ranks $(r_1,r_2, \ldots,r_k)$ are set a priori, the goal is now to find the optimal projection matrices
\begin{equation}
	\left\{\mat{U}^*\right\} = \argmax_{\{\mat{U}\}}\phi
\end{equation}
which is solved through the HODA algorithm.
To start, $\mat{U}_k$ are initialized to orthogonal matrices, e.g. as random
orthonormal matrices, by a per mode Singular Value Decomposition (SVD),
or as the partial HOSVD of all stacked tensors in the dataset.
At each iteration, the algorithm loops through the modes and fixes all
projections but $\mat{U}_k$ corresponding to mode $k$.
It then finds a partial latent tensor
\begin{equation}
	\ten{G}_{-k}=\ten{X}\mmprs{\mat{U}}{k}
\end{equation}
Subsequently, a new projection matrix $\mat{V}_k$ can be found analogous to Linear
Discriminant Analysis by constructing the within-class scatter matrix
\begin{equation}
	\mat{S}_{-k,\text{w}} = \sum_n^N\tilde{\mat{G}}_{-k,k}(n)\cdot\tilde{\mat{G}}_{-k,k}^\intercal(n)
\end{equation}
with $\tilde{\ten{G}}_{-k}(n) = \ten{G}_{-k}(n) - \bar{\ten{G}}_{-k}(c_n)$,
and the between-class scatter matrix
\begin{equation}
	\mat{S}_{-k,\text{b}} =
	\sum_c^CN_c\tilde{\bar{\mat{G}}}_{-k,k}(c)\cdot\tilde{\bar{\mat{G}}}_{-k,k}^\intercal(c)
\end{equation}
with $\tilde{\bar{\ten{G}}}_{-k}(c) = \bar{\ten{G}}_{-k}(c) - \bar{\bar{\ten{G}}}_{-k}$,
and solving for the $r_k$ leading eigenvectors in the eigenvalue problem
\begin{equation}
	\mat{S}_{-k,\text{b}}-\varphi_k\mat{S}_{-k,\text{w}} =
	\mat{V}_k\mat{\Lambda}\mat{V}_k^\intercal
\end{equation}
with $\varphi_k=\Tr\left(\mat{U}_k^\intercal\mat{S}_{-k,\text{b}}\mat{U}_k\right)/\Tr\left(\mat{U}_k^\intercal\mat{S}_{-k,\text{w}}\mat{U}_k\right)$
using the $\mat{U}_k$ obtained in the previous iteration.
Finally, the orthogonal transformation invariant projections $\mat{U}_k$
are obtained by calculating the
per-mode total scatter matrices
\begin{equation}
  \mat{S}_{k,\text{t}} = \sum_n^N\mat{X}_k(n)\cdot\mat{X}_k^\intercal(n)
\end{equation}
and finding the $r_k$ leading eigenvectors of
\begin{equation}
	\mat{V}_k\mat{V}_k^\intercal\mat{S}_{k,\text{t}}\mat{V}_k\mat{V}_k^\intercal
	= \mat{U}_k\mat{\Lambda}\mat{U}_k^\intercal
\end{equation}
at each iteration~\cite{Wang2007}.
%and solving the generalized eigenvalue problem
%\begin{equation}
%  = \mat{\lambda}\mat{S}_{-k,\text{w}}\mat{U}_k
%\end{equation}
The iterative process halts after a fixed amount or iterations, or when the
update of each $\mat{U}_k$ is lower than a predetermined threshold.
The full HODA procedure is summarized in Algorithm~\ref{alg:HODA}.
\begin{algorithm}
  \caption[The \ac{hoda} backward solution.]{The \ac{hoda} backward solution.}
	\label{alg:HODA}
	\input{chapters/bttda/alg_HODA.tex}
\end{algorithm}

@@

\subsection{Block-Term Tensor Discriminant Analysis}
After defining the forward model, we can construct our proposed block-term
tensor model.
Assuming the latent tensors $\ten{G}$
obtained by the backward projection of HODA do not achieve perfect
class separation, the error term $\ten{\mathbfcal{E}}$ in
Equation~\ref{eq:HODA-forward} should still contain some discriminative
information, which can be exploited to improve classifier
performance.
Useful features can still be extracted from $\ten{\mathbfcal{eps}} = \ten{X} -
\hat{\ten{X}}$, so it is further projected onto another core tensor
$\ten{G}^{(2)}$ (assuming $\ten{G}$ as $\ten{G}^{(1)}).$

@@

\subsection{Block contribution}
\todo{Discuss number of features}
To analyze the contribution of extra feature blocks extracted by {BTTDA} over
the first one found by HODA we pick n ERP ($K=2$) dataset
(BNCI2014-008) and an MI ($K=3$) dataset (BNCI2014-001).
We report within-session ROC-AUC scores for training, validation and test data as a function
of the number of blocks increases, Figure~\ref{fig:blocks}.
Training and validation folds were taken from the model selection procedure.
Additionally, the Normalized Mean Squared Error (NMSE) is reported for the
reconstructed from the truncated BTTDA decomposition
$\textstyle{\hat{\ten{X}}^{(B)}=\sum_b^B\ten{G}^{(b)}\mmpr{\mat{U}^{(b)}}}$.
NMSE is calculated as
\begin{equation}
  \nmse\left(\ten{X}, \hat{\ten{X}}^{(B)}\right) =
	\frac{\sum_n^N\left\lVert\ten{X}(n)-\ten{X}_\text{rec}^{(B)}(n)\right\rVert_F^2}
	{\sum_n^N\left\lVert\ten{X}(n)\right\rVert_F^2}
\end{equation}
\begin{figure*}[t]
	\input{figures/bttda/blocks.pgf}
  \caption[Analysis of NMSE and classification score per block.]{%
    Normalized Mean Square Error (NMSE) (left), and difference in
    classification score for the training and validation for the ERP datasets
    (top row) and the MI dataset (bottom row)	of the greedy model selection
    procedure (right), as a function of the number of BTTDA blocks $b$.
		While NMSE monotonically decreases for the evaluated datasets, better class
		separation will be achieved, but eventually overfitting occurs and validation
		and test scores will drop, or plateau due to feature selection.
	}
	\label{fig:blocks}
\end{figure*}
\todo{format figure}

@@

Additionally, the forward modeling step inherent to the BTTDA results
in an interpretable model, since we can use the activation patterns or the
forward projection to inspect the neural patterns corresponding to the
relevant discriminatory information at each block~\cite{Haufe2014}.
Figure~\ref{fig:forward} shows the activation patterns
of two blocks obtained from the BNCI2014-008 dataset as well as the forward
projection of the difference between the averages of the mean latent tensor per
class (\emph{contrasts}) after forward projection.
While the weights of the backward projection are
uninterpretable~\cite{Haufe2014},
the activation patterns and contrasts after forward projection clearly show
that ERP components can be recognized and separated into different
BTTDA blocks.
\begin{figure*}[t]
	\includegraphics[width=\linewidth]{figures/bttda/forward_block-0.png}
	\includegraphics[width=\linewidth]{figures/bttda/forward_block-1.png}
  \caption[Extracted \ac{bttda} activation patterns.]{%
    Spatial (left two columns) and temporal (middle column) activation patterns and
		condition contrasts (right column) obtained after forward projection of the latent
		features for 2 blocks (row 1 and row 2) of rank $(2,2)$ of BTTDA
		fit on the full dataset BNCI2014-008.
		The separate blocks approximately model different ERP
		components.}
	\label{fig:forward}
\end{figure*}
Given informed or correctly tuned hyperparameters, this method could be used to
e.g. separate ERP components or neural processes based on the task-related
information in the class labels.

>>>>

\subsection{Higher Order Discriminant Analysis}
Higher Order Discriminant Analysis (HODA)~\cite{Phan2010} is a
supervised tensor-based feature extraction technique.
For a set of $N$ tensors of order $K$
$\left\{\ten{X}(n)\in\mathbb{R}^{D_1\times D_2 \times \cdots \times
		D_K}\right\}_n^N$, HODA finds projection matrices $\mat{U_k}$ for each mode $k$
that project a given $\ten{X}$ to a latent tensor
$\ten{G}\in\mathbb{R}^{r_1\times r_2\times\cdots\times r_K}$, usually with lower
dimensionality $(r_1\leq D_1,r_2\leq D_2,\ldots,r_K\leq D_K)$ using
tensor-matrix mode products
\begin{equation}
	\ten{G}  = \ten{X}\mmpr{\mat{U}}
	\label{eq:HODA-backward}
\end{equation}
visualized in Figure~\ref{fig:HODA-backward}.
\begin{figure}[t]
	\centering
	\input{figures/bttda/HODA_backward.tikz.tex}
  \caption[The \acs{hoda} backward projection.]{%
    A visualization of the multilinear projection obtained by Higher Order
    Discriminant Analysis (HODA) applied to a third-order tensor
    sample $\ten{X}$ with shape $(D_1,D_2, D_3)$.
		HODA finds projection matrices $\mat{U}_k$ such that maximal
		discriminability between classes is achieved in the projected latent tensors
		$\ten{G}$ with reduced dimensionality $(r_1,r_2,r_3)$.}
	\label{fig:HODA-backward}
\end{figure}
Analogous to the HOSVD, HODA is a dimensionality
reduction decomposition that results in a dense latent tensor $\ten{G}$, and
imposes an orthogonality constraint on $\mat{U}_k$ to ensure uniqueness.
However, while for the HOSVD decomposition the projection matrices
are chosen to minimize the reconstruction error, the projection matrices
$\mat{U}_k$ of HODA are optimized for maximal discriminability of tensors
$\ten{G}(n)$ belonging to classes with labels $c_n$.
This is a desirable property in a classification setting where samples are
high-dimensional tensors.

HODA optimizes discriminability in the Fisher sense, by optimizing the
Fisher ratio $\phi$ between the latent tensors $\ten{G}(n)$
\begin{equation}
	\phi = \frac{\sum_c^CN_c\left\lVert\bar{\ten{G}}(c)-\bar{\bar{\ten{G}}}\right\rVert_F^2}
	{\sum_n^N\left\lVert\ten{G}(n)-\bar{\ten{G}}(c_n)\right\rVert_F^2}
	\label{eq:fisher}
\end{equation}
for $C$ classes with each $N_c$ samples. $\bar{\ten{G}}(c)$ is the mean of
latent
tensors of class $c$, and $\bar{\bar{\ten{G}}}$ the mean of
these class mean latent tensors.
If the ranks $(r_1,r_2, \ldots,r_k)$ are set a priori, the goal is now to find the optimal projection matrices
\begin{equation}
	\left\{\mat{U}^*\right\} = \argmax_{\{\mat{U}\}}\phi
\end{equation}
which is solved through the HODA algorithm.
To start, $\mat{U}_k$ are initialized to orthogonal matrices, e.g. as random
orthonormal matrices, by a per mode Singular Value Decomposition (SVD),
or as the partial HOSVD of all stacked tensors in the dataset.
At each iteration, the algorithm loops through the modes and fixes all
projections but $\mat{U}_k$ corresponding to mode $k$.
It then finds a partial latent tensor
\begin{equation}
	\ten{G}_{-k}=\ten{X}\mmprs{\mat{U}}{k}
\end{equation}
Subsequently, a new projection matrix $\mat{V}_k$ can be found analogous to Linear
Discriminant Analysis by constructing the within-class scatter matrix
\begin{equation}
	\mat{S}_{-k,\text{w}} = \sum_n^N\tilde{\mat{G}}_{-k,k}(n)\cdot\tilde{\mat{G}}_{-k,k}^\intercal(n)
\end{equation}
with $\tilde{\ten{G}}_{-k}(n) = \ten{G}_{-k}(n) - \bar{\ten{G}}_{-k}(c_n)$,
and the between-class scatter matrix
\begin{equation}
	\mat{S}_{-k,\text{b}} =
	\sum_c^CN_c\tilde{\bar{\mat{G}}}_{-k,k}(c)\cdot\tilde{\bar{\mat{G}}}_{-k,k}^\intercal(c)
\end{equation}
with $\tilde{\bar{\ten{G}}}_{-k}(c) = \bar{\ten{G}}_{-k}(c) - \bar{\bar{\ten{G}}}_{-k}$,
and solving for the $r_k$ leading eigenvectors in the eigenvalue problem
\begin{equation}
	\mat{S}_{-k,\text{b}}-\varphi_k\mat{S}_{-k,\text{w}} =
	\mat{V}_k\mat{\Lambda}\mat{V}_k^\intercal
\end{equation}
with $\varphi_k=\Tr\left(\mat{U}_k^\intercal\mat{S}_{-k,\text{b}}\mat{U}_k\right)/\Tr\left(\mat{U}_k^\intercal\mat{S}_{-k,\text{w}}\mat{U}_k\right)$
using the $\mat{U}_k$ obtained in the previous iteration.
Finally, the orthogonal transformation invariant projections $\mat{U}_k$
are obtained by calculating the
per-mode total scatter matrices
\begin{equation}
  \mat{S}_{k,\text{t}} = \sum_n^N\mat{X}_k(n)\cdot\mat{X}_k^\intercal(n)
\end{equation}
and finding the $r_k$ leading eigenvectors of
\begin{equation}
	\mat{V}_k\mat{V}_k^\intercal\mat{S}_{k,\text{t}}\mat{V}_k\mat{V}_k^\intercal
	= \mat{U}_k\mat{\Lambda}\mat{U}_k^\intercal
\end{equation}
at each iteration~\cite{Wang2007}.
%and solving the generalized eigenvalue problem
%\begin{equation}
%  = \mat{\lambda}\mat{S}_{-k,\text{w}}\mat{U}_k
%\end{equation}
The iterative process halts after a fixed amount or iterations, or when the
update of each $\mat{U}_k$ is lower than a predetermined threshold.
The full HODA procedure is summarized in Algorithm~\ref{alg:HODA}.
\begin{algorithm}
  \caption[The \acs{hoda} backward solution.]{The \ac{hoda} backward solution.}
	\label{alg:HODA}
	\input{chapters/bttda/alg_HODA.tex}
\end{algorithm}

@@

\subsection{Block-Term Tensor Discriminant Analysis}
After defining the forward model, we can construct our proposed block-term
tensor model.
Assuming the latent tensors $\ten{G}$
obtained by the backward projection of HODA do not achieve perfect
class separation, the error term $\ten{\mathbfcal{E}}$ in
Equation~\ref{eq:HODA-forward} should still contain some discriminative
information, which can be exploited to improve classifier
performance.
Useful features can still be extracted from $\ten{\mathbfcal{E}} = \ten{X} -
\hat{\ten{X}}$, so it is further projected onto another core tensor
$\ten{G}^{(2)}$ (assuming $\ten{G}$ as $\ten{G}^{(1)}).$

@@

\subsection{Block contribution}
To analyze the contribution of extra feature blocks extracted by {BTTDA} over
the first one found by HODA we pick n ERP ($K=2$) dataset
(BNCI2014-008) and an MI ($K=3$) dataset (BNCI2014-001).
We report within-session ROC-AUC scores for training, validation and test data as a function
of the number of blocks increases, Figure~\ref{fig:blocks}.
Training and validation folds were taken from the model selection procedure.
Additionally, the Normalized Mean Squared Error (NMSE) is reported for the
reconstructed from the truncated BTTDA decomposition
$\textstyle{\hat{\ten{X}}^{(B)}=\sum_b^B\ten{G}^{(b)}\mmpr{\mat{U}^{(b)}}}$.
NMSE is calculated as
\begin{equation}
  \nmse\left(\ten{X}, \hat{\ten{X}}^{(B)}\right) =
	\frac{\sum_n^N\left\lVert\ten{X}(n)-\ten{X}_\text{rec}^{(B)}(n)\right\rVert_F^2}
	{\sum_n^N\left\lVert\ten{X}(n)\right\rVert_F^2}
\end{equation}
\begin{figure*}[t]
	\input{figures/bttda/blocks.pgf}
  \caption[Analysis of NMSE and classification score per block.]{%
    Normalized Mean Square Error (NMSE) (left), and difference in
    classification score for the training and validation for the ERP datasets
    (top row) and the MI dataset (bottom row)	of the greedy model selection
    procedure (right), as a function of the number of BTTDA blocks $b$.
		While NMSE monotonically decreases for the evaluated datasets, better class
		separation will be achieved, but eventually overfitting occurs and validation
		and test scores will drop, or plateau due to feature selection.
	}
	\label{fig:blocks}
\end{figure*}
\todo{format figure}

@@

Additionally, the forward modeling step inherent to the BTTDA results
in an interpretable model, since we can use the activation patterns or the
forward projection to inspect the neural patterns corresponding to the
relevant discriminatory information at each block~\cite{Haufe2014}.
Figure~\ref{fig:forward} shows the activation patterns
of two blocks obtained from the BNCI2014-008 dataset as well as the forward
projection of the difference between the averages of the mean latent tensor per
class (\emph{contrasts}) after forward projection.
While the weights of the backward projection are
uninterpretable~\cite{Haufe2014},
the activation patterns and contrasts after forward projection clearly show
that ERP components can be recognized and separated into different
BTTDA blocks.
\begin{figure*}[t]
	\includegraphics[width=\linewidth]{figures/bttda/forward_block-0.png}
	\includegraphics[width=\linewidth]{figures/bttda/forward_block-1.png}
  \caption[Extracted \acs{bttda} activation patterns.]{%
    Spatial (left two columns) and temporal (middle column) activation patterns and
		condition contrasts (right column) obtained after forward projection of the latent
		features for 2 blocks (row 1 and row 2) of rank $(2,2)$ of BTTDA
		fit on the full dataset BNCI2014-008.
		The separate blocks approximately model different ERP
		components.}
	\label{fig:forward}
\end{figure*}
Given informed or correctly tuned hyperparameters, this method could be used to
e.g. separate ERP components or neural processes based on the task-related
information in the class labels.
